% !TEX root = ../master.tex
\section{Semantics}
The following section presents the set of formal rules that define how constraints are extracted from a program.
To do this we define the environments we require for our rules and then the transition systems in which the rules exist.

\myparagraph{Label environment}
The label environment is called $\iEnvL$ and is defined by:
\[
  \iEnvL = \iVar \cup \iFun \cup \ia \cup \ib \cup \ifb \rightharpoonup \mathbb{L}
\]

where $\ia$ is the authority label, $\ib$ the block label (also known as the program counter), and $\ifb$ the function block label.
The latter is used to refer to the label of a function, when the body of that function is evaluated.

$\mathbb{L}$ is the set of all possible labels, with the six types of labels: \\

\begin{tabular}{rl}
  \textbf{Policy labels}    & Like those described earlier \\
  \textbf{Variable labels}  & Denoted $\vlbl{x}$ \\
  \textbf{Constant labels}  & Denoted $\clbl{p}$ \\
  \textbf{Lower bound}      & Denoted $\bot$ \\
  \textbf{Upper bound}      & Denoted $\top$ \\
  \textbf{Join/meet labels} & Combinations of the above-mentioned labels
\end{tabular}

\myparagraph{Function parameters environment}
An environment is used to store the collection of parameters and their labels, such that this information can be applied when calling a function.
The function parameters environment is called $\iEnvP$ and is defined by:
\[
  %\iEnvP = \iFun \rightharpoonup \mathbf{P}(\iVar \times \mathbb{L})
  \iEnvP = \iFun \rightharpoonup (\iVar \times \mathbb{L})_1 \cup (\iVar \times \mathbb{L})_2 \cup \dots \cup (\iVar \times \mathbb{L})_n
\]

\myparagraph{Constraint set}
In order to capture the set of constraints in a program, we require a definition for a set of constraints.
We use $\iCstr$ to denote a set of constraints where each $c \in \iCstr$ is on the form:
\[ L_1 \sqsubseteq L_2 \; | \; L_1, L_2 \in \mathbb{L} \]

\myparagraph{Program transition system}
For the syntactical category $\iProg$ we provide the following transition system.
The system handles the declaration of global variables and functions.
The $\iCstr$ defined in the system output represents the collection of all constraints in that program.
\mikkelin{Hvordan definerer vi et ``entry-point''?}

$(\Gamma_P, \rightarrow_P, T_P)$, with
\begin{align*}
  \Gamma_P  & = (\iProg \times \iEnvL \times \iEnvP) \cup (\iEnvL \times \iCstr \times \iEnvP) \\
  T_P       & = (\iEnvL \times \iCstr \times \iEnvP)
\end{align*}

\myparagraph{Statement transition system}
For the syntactical category $\iStm$ we provide the following transition system.
The system evaluates statements, effectively extracting any constraints therein.
As functions can only be declared in the global scope, the $\iEnvP$ is not directly a part of the transition system.

$(\Gamma_S, \rightarrow_S, T_S)$, with
\begin{align*}
  \Gamma_S & = (\iStm \times \iEnvL) \cup (\iEnvL \times \iCstr) \\
  T_S & = (\iEnvL \times \iCstr)
\end{align*}

Where transitions are on the form:
\[ \ienvP \vdash \argpart{S, \ienvL}{S}{(\ienvL, \icstr)} \]

\myparagraph{Label transition system}
For the syntactical category $\iLbl$ and $\iPol$ we provide the following transition system.
The system translates the literals used for label declarations into actual labels.

Note the use of $\varepsilon$, which allows for missing label declarations to be passed through the transition system.

$(\Gamma_L, \rightarrow_L, T_L)$, with
\begin{align*}
  \Gamma_L & = \iLbl \cup \iPol \cup \mathbb{L} \cup \varepsilon \\
  T_L & = \mathbb{L} \cup \varepsilon
\end{align*}

Where transitions are one of the following two forms:
\[ \ienvL \vdash l \rightarrow_L L \]
\[ \ienvL \vdash l \rightarrow_L \varepsilon \]

\myparagraph{Expression transition system}
For the syntactical category $\iExp$ we provide the following transition system.
The system evaluates expressions and returns any constraints in an expression along with the label value of that expression.

$(\Gamma_e, \rightarrow_e, T_e)$, with
\begin{align*}
  \Gamma_e =  & \;      \iExp \cup (\iCstr \times \mathbb{L}) \\
  T_e =       & \; (\iCstr \times \mathbb{L})
\end{align*}

Where transitions are of the form $\ienvL, \ienvP \vdash e \rightarrow_e (\icstr, L)$.

\subsection{Label declarations}
The first set of rules will handle the translation of all label declarations.
Note that these rules do not provide an $\mathbb{L}$ value for an undeclared label.
As the label value for an undeclared label depends on its context, $\varepsilon$ is simply passed through on evaluation.

\begin{trules}
  \ctrule{LABEL}
        {\ienvL \vdash \tk{\{\{} pol \tk{\}\}} \rightarrow_L L}
        {\ienvL \vdash pol \rightarrow_L L}
        {}
  \trule{LABEL-EMPTY}
        {\ienvL \vdash \varepsilon \rightarrow_L \varepsilon}
        {}
  \ctrule{POLICY-COMP}
        {\ienvL \vdash pol_1 \tk ; pol_2 \rightarrow_L L_1 \sqcup L_2}
        {\ienvL \vdash pol_1 \rightarrow_L L_1 \quad pol_2 \rightarrow_L L_2}
        {}
  \trule{POLICY-VAR}
        {\ienvL \vdash x \rightarrow_L L}
        {where $\ienvL \, x = L$}
  \trule{POLICY-PRINCIPALS}
        {\ienvL \vdash p_0 \tk{->} p_1 \tk , p_2 \tk , \dots \tk , p_k \rightarrow_L L}
        {where $p_i' = \mathcal{P}(p_i)$, for $0 \leq i \leq k$ \\
          and $L = \{p_0' \rightarrow p_1', p_2', \dots, p_k'\}$}
\end{trules}

\subsection{Program}
The rules for programs can be split in two groups; the two that are simple and the one that is not.
Note however that variable declaration is only this simple because the actual evaluation is deferred to the $\rightarrow_S$ transition system, which is detailed in \Cref{semantic:statements}.
\begin{trules}
\ctrule {PROG-COMP}
        {\argpart{P_1 \; P_2, \ienvL, \ienvP}{P}{(\ienvL_2, \icstr_1 \cup \icstr_2, \ienvP_2)}}
        {\begin{gathered}
          \argpart{P_1, \ienvL, \ienvP}{P}{(\ienvL_1, \icstr_1, \ienvP_1)} \\
          \argpart{P_2, \ienvL_1, \ienvP_1}{P}{(\ienvL_2, \icstr_2, \ienvP_2)}
        \end{gathered}}
        {}
\ctrule {PROG-VAR-DECL}
        {\argpart {D_V} {P} {(\ienvL_1, \icstr, \ienvP)} }
        {\ienvP \vdash \argpart{D_V, \ienvL}{S}{(\ienvL_1, \icstr)}}
        {}
\end{trules}

\mikkelin{Vi har lavet en let blanding af $'$ og sub-script i den her regel.
Det er endnu ikke afklaret hvad vores løsning bliver.}
Function declaration (along with function call) can easily be considered the most complex semantic rule(s).
Specifically the many different label environments in function declarations can render to rule complex.

In a function declarations return label $l_f$ it is possible to reference the parameters of that function.
To achieve this the $\ienvL'$ environment is constructed from the parameters.

When evaluating the statements of the function a reference to the functions return label is required; this is included in $\ienvL'''$ along with a label reference for the declared function itself.
The latter allows for recursion.

Finally $\ienvL''$ represents the label environment exposed to other functions that might call the function.

\ctrule {FUNC-DECL}
        {\argpart{T_f \, l_f \, f \tk ( T_1 \, l_1 \, x_1 \tk , T_2 \, l_2 \, x_2 \tk , \dots \tk , T_n \, l_n \, x_n \tk ) S, \ienvL, \ienvP}{P}{(\ienvL'', \icstr, \ienvP')}}
        {\begin{gathered}
          \ienvL \vdash l_1 \rightarrow_L L_1 \quad l_2 \rightarrow_L L_2 \quad \dots \quad l_n \rightarrow_L L_n \\
          \ienvL' \vdash l_f \rightarrow_L L_f \quad \ienvP' \vdash \argpart{S, \ienvL'''}{S}{(\ienvL'''', \icstr)}
        \end{gathered}}
        {where $L_{x_i} = \begin{cases}
          \clbl{x_i} & \quad \text{if } L_i = \varepsilon \\
          L_i & \quad otherwise
        \end{cases}\quad \text{for } 1 \leq i \leq n$ \\
          and $\ienvL' = \ienvL[x_1 \mapsto L_{x_1}, x_2 \mapsto L_{x_2}, \dots, x_n \mapsto L_{x_n}]$ \\
          and $L_f' = \begin{cases}
            \mathlarger\sqcupl_{i = 1}^n L_{x_i} & \quad \text{if } L_f = \varepsilon \\
            L_f & \quad \text{otherwise}
        \end{cases}$ \\
          and $\ienvL'' = \ienvL[f \mapsto L_f']$ \\
          and $\ienvL''' = \ienvL'[f \mapsto L_f', \ifb \mapsto L_f']$\\
          and $\ienvP' = \ienvP[f \mapsto ((L_{x_1}, x_1), (L_{x_2}, x_2), \dots, (L_{x_n}, x_n))]$}

\subsection{Statements}\label{semantic:statements}
For statement evaluation we first cover a few simple rules.
The expression statement allows us to perform function calls where we are not interested in the return value.

\begin{trules}
  \trule{EMPTY}
        {\ienvP \vdash \argpart{\varepsilon, \ienvL}{S}{(\ienvL, \emptyset)}}
        {}
  \ctrule{EXPR}
        {\ienvP \vdash \argpart{e, \ienvL}{S}{(\ienvL, \icstr)}}
        {\ienvL, \ienvP \vdash e \rightarrow_e (\icstr, L_e)}
        {}
  \ctrule{COMP}
        {\ienvP \vdash \argpart{S_1 \; S_2, \ienvL}{S}{(\ienvL_2, \icstr_1 \cup \icstr_2)}}
        {\ienvP \vdash \argpart{S_1, \ienvL}{S}{(\ienvL_1, \icstr_1)} \quad
          \ienvP \vdash \argpart{S_2, \ienvL_1}{S}{(\ienvL_2, \icstr_2)}}
        {}
\end{trules}

\myparagraph{Variable declarations}
New labels are introduced when declaring variables.
A variable declaration without initialization uses $\bot$ as the label value for the missing initialization value.

\begin{trules}
  \ctrule{VAR-DECL-EXPR}
        {\ienvP \vdash \argpart {T \, l \, x \tk = e, \ienvL} {S} {(\ienvL[x \mapsto L_x], \icstr \cup \{ c \})} }
        {\ienvL \vdash l \rightarrow L \quad
          \ienvL, \ienvP \vdash e \rightarrow_e (\icstr, L_e)}
        {where $L_x = \begin{cases}
            \vlbl x & \quad \text{if } L = \varepsilon \\
            L & \quad \text{otherwise}
          \end{cases}$ \\
          and $L_\ib = \ienvL \, \ib$ \\
          and $c = " \, L_e \sqcup L_\ib \sqsubseteq L_x \, "$ }
  \ctrule{VAR-DECL}
        {\ienvP \vdash \argpart {T \, l \, x, \ienvL} {S} {(\ienvL[x \mapsto L_x], \icstr \cup \{ c \})} }
        {\ienvL \vdash l \rightarrow L}
        {where $L_x = \begin{cases}
            \vlbl x & \quad \text{if } L = \varepsilon \\
            L & \quad \text{otherwise}
          \end{cases}$ \\
          and $L_\ib = \ienvL \, \ib$ \\
          and $c = "\bot \sqcup L_\ib \sqsubseteq L_x"$ }
\end{trules}

\myparagraph{Control flow}
The control flow constructs (while and if) define constraints that will allow inference to determine their label basic blocks.
These constraints allows for propagation of label constraints and are effectively the implementation of implicit flows.

\begin{trules}
  \ctrule{WHILE}
        {\ienvP \vdash \argpart{\tk{while} \tk ( e \tk ) S, \ienvL}{S}{(\ienvL, \icstr_1 \cup \icstr_2 \cup \{ c \})}}
        {\ienvL_1, \ienvP \vdash e \rightarrow_e (\icstr_1, L_e) \quad
          \ienvP \vdash \argpart{S, \ienvL_1}{S}{(\ienvL_2, \icstr_2)}}
        {where $L_\ib = \ienvL \, \ib$ \\
          and $L_w = next$ \\
          and $\ienvL_1 = \ienvL[\ib \mapsto L_w]$ \\
          and $c = " \, L_e \sqcup L_\ib \sqsubseteq L_w \, "$}
  \ctrule{IF}
        {\ienvP \vdash \argpart{\tk{if} \tk ( e \tk ) S, \ienvL}{S}{(\ienvL, \icstr_1 \cup \icstr_2 \cup \{ c \})}}
        {\ienvL, \ienvP \vdash e \rightarrow_e (\icstr_1, L_e) \quad
          \ienvP \vdash \argpart{S, \ienvL_1}{S}{(\ienvL_2, \icstr_2)}}
        {where $L_\ib = \ienvL \, \ib$ \\
          and $L_{if} = next$ \\
          and $\ienvL_1 = \ienvL[\ib \mapsto L_{if}]$ \\
          and $c = " \, L_e \sqcup L_\ib \sqsubseteq L_{if} \, "$}
  \ctrule{IF-ELSE}
        {\ienvP \vdash \argpart{\tk{if} \tk ( e \tk ) S_1 \tk{else} S_2, \ienvL}{S}{(\ienvL, \icstr_1 \cup \icstr_2 \cup \icstr_3 \cup \{ c \})}}
        {\begin{gathered}
          \ienvL, \ienvP \vdash e \rightarrow_e (\icstr_1, L_e) \\
          \ienvP \vdash \argpart{S_1, \ienvL_1}{S}{(\ienvL_2, \icstr_2)} \\
            \ienvP \vdash \argpart{S_2, \ienvL_1}{S}{(\ienvL_3, \icstr_3)}
        \end{gathered}}
          {where $L_\ib = \ienvL \, \ib$ \\
            and $L_{if} = next$ \\
            and $\ienvL_1 = \ienvL[\ib \mapsto L_{if}]$ \\
            and $c = " \, L_e \sqcup L_\ib \sqsubseteq L_{if} \, "$}
\end{trules}

\myparagraph{Assignments and return statements}
It should be noticed that the constraints constructed for declarations, assignments and return statements have a similar structure.
This structure represents the similarity in the feature they are handling, be it assignment to a variables value or to the return value of a function.

\begin{trules}
  \ctrule{ASS}
        {\ienvP \vdash \argpart{x \tk = e, \ienvL}{S}{(\ienvL, \icstr \cup \{ c \})}}
        {\ienvL, \ienvP \vdash e \rightarrow_e (\icstr, L_e)}
        {where $L_\ib = \ienvL \, \ib$ \\
          and $c = " \, L_e \sqcup L_\ib \sqsubseteq L_x \, "$ }
  \ctrule{RET-EXPR}
        {\ienvP \vdash \argpart{\tk{return} e, \ienvL}{S}{(\ienvL, \icstr \cup \{ c \})}}
        {\ienvL, \ienvP \vdash e \rightarrow_e (\icstr, L_e)}
        {where $L_\ib = \ienvL \, \ib$ \\
          and $L_\ifb = \ienvL \, \ifb$ \\
          and $c = " \, L_e \sqcup L_\ib \sqsubseteq L_\ifb \, "$}
  \trule{RET}
        {\ienvP \vdash \argpart{\tk{return}, \ienvL}{S}{(\ienvL, \emptyset)}}
        {}
\end{trules}

\myparagraph{Acts for statements}
Finally the acts for statements have identical rules.
This is because the difference between the two only exists at runtime.
The $L_a$ label represents the effective authority on execution.
This is stored in the label environment to be used for implicit declassification when evaluating expressions.

\begin{trules}
  \ctrule{ACTS-THIS}
        {\ienvP \vdash \argpart{\tk{this} \tk{-{}->?} p_1 \tk , p_2 \tk , \dots \tk , p_k \; S, \ienvL}{S}{(\ienvL, \icstr)}}
        {\ienvP \vdash \argpart{S, \ienvL[\ia \mapsto L_\ia']}{S}{(\ienvL_1, \icstr)}}
        {where $L_\ia = \ienvL \, \ia$ \\[0.5em]
          and $L_\ia' = L_\ia \sqcup \mathlarger\sqcupl_{i=1}^k \, \{ p_i \rightarrow \emptyset \}$}
    \ctrule{ACTS-CALLER}
        {\ienvP \vdash \argpart{\tk{caller} \tk{-{}->?} p_1 \tk , p_2 \tk , \dots \tk , p_k \; S, \ienvL}{S}{(\ienvL, \icstr)}}
        {\ienvP \vdash \argpart{S, \ienvL[\ia \mapsto L_\ia']}{S}{(\ienvL_1, \icstr)}}
        {where $L_\ia = \ienvL \, \ia$ \\[0.5em]
          and $L_\ia' = L_\ia \sqcup \mathlarger\sqcupl_{i=1}^k \, \{ p_i \rightarrow \emptyset \}$}
\end{trules}

\subsection{Expressions}
Most rules for expression evaluation produce no constraints and have a simple definition in terms of the label they return.
\begin{trules}
  \trule  {VAR}
          {\ienvL, \ienvP \vdash x \rightarrow_e (\emptyset, L)}
          {where $\ienvL \, x = L$}
  \trule  {CONST}
          {\ienvL, \ienvP \vdash K \rightarrow_e (\emptyset, \bot)}
          {}
  \ctrule {BINARY-OP}
          {\ienvL, \ienvP \vdash e_1 \; op_b \; e_2 \rightarrow_e (\icstr_1 \cup \icstr_2, L_1 \sqcup L_2)}
          {\ienvL, \ienvP \vdash e_1 \rightarrow_e (\icstr_1, L_1) \;\; \ienvL, \ienvP \vdash e_2 \rightarrow_e (\icstr_2, L_2)}
          {}
  \ctrule {UNARY-OP}
          {\ienvL, \ienvP \vdash op_u \; e \rightarrow_e (\icstr, L)}
          {\ienvL, \ienvP \vdash e \rightarrow_e (\icstr, L)}
          {}
  \ctrule {PAREN}
          {\ienvL, \ienvP \vdash \tk ( e \tk ) \rightarrow_e (\icstr, L)}
          {\ienvL, \ienvP \vdash e \rightarrow_e (\icstr, L)}
          {}
\end{trules}

\myparagraph{Declassification}
The two rules for declassification shows some of the simplicity that is part of inference.
Either an explicit label is defined for declassification or one (represented by the variable label $L_d$) will be inferred.
Asides from that the rules are exactly the same.

\begin{trules}
  \ctrule{DEC}
        {\ienvL, \ienvP \vdash \tk{<|} e \tk{|>} \rightarrow_e (\icstr \cup \{ c \}, L_d)}
        {\ienvL, \ienvP \vdash e \rightarrow_e (\icstr, L_e)}
        {where $L_\ia = \ienvL \, \ia$ \\
          and $L_{d} = next$ \\
          and $c = " \, L_e \sqsubseteq L_d \sqcup L_\ia \, "$}
  \ctrule{DEC-EXPL}
        {\ienvL, \ienvP \vdash \tk{<|} e \tk , l \tk{|>} \rightarrow_e (\icstr \cup \{ c \}, L)}
        {\ienvL \vdash l \rightarrow L \quad \ienvL, \ienvP \vdash e \rightarrow_e (\icstr, L_e)}
        {where $L_\ia = \ienvL \, \ia$ \\
          and $c = " \, L_e \sqsubseteq L \sqcup L_\ia \, "$}
\end{trules}

\subsection{Function call}
The final type of expression is a function call.
There are two rules regarding function calls.
One for functions that have label definitions and one for those that do not.

To represent function calls for functions with labels we require an additional transition system that can aid in determining the return label of a function call.
The systems is as follows:

$(\mathbb{L}, \rightarrow_F, \mathbb{L})$,

where transitions are of the form $\ienvL \vdash L_1 \rightarrow_F L_2$.
%policy, var, const, low, up, join, meet

Below are the rules for each type of label.
The effect of applying this transition system to a label is that its constant labels are replaced with other labels.
We can use this to replace the constant labels in a functions return label with the labels of its arguments.

\begin{trules}
  \trule {POLICY}
         {\ienvL \vdash \literal{p_0 \rightarrow p_1, p_2, \dots, p_n} \rightarrow_F \literal{p_0 \rightarrow p_1, p_2, \dots, p_n}}
         {}
  \trule {CONSTANT}
         {\ienvL \vdash \clbl x \rightarrow_F \ienvL \, x}
         {}
  \trule {VARIABLE}
         {\ienvL \vdash \vlbl x \rightarrow_F \vlbl x}
         {}
  \ctrule {JOIN}
         {\ienvL \vdash \literal{L_1 \sqcup L_2} \rightarrow_F \literal{L_1' \sqcup L_2'}}
         {\ienvL \vdash L_1 \rightarrow_F L_1' \quad L_2 \rightarrow_F L_2'}
         {}
  \ctrule {MEET}
         {\ienvL \vdash \literal{L_1 \sqcap L_2} \rightarrow_F \literal{L_1' \sqcap L_2'}}
         {\ienvL \vdash L_1 \rightarrow_F L_1' \quad L_2 \rightarrow_F L_2'}
         {}
  \trule {LOWER-BOUND}
         {\ienvL \vdash \bot \rightarrow_F \bot}
         {}
  \trule {UPPER-BOUND}
         {\ienvL \vdash \top \rightarrow_F \top}
         {}
\end{trules}

Using the system above we can handle function calls.
Though only functions that are defined using out syntax (with explicit or implicit labeling) will make use of it.
The label value of a function call for functions that do not have a return label will evaluate to the join of the arguments labels.
This allows for the use of functions that have not been written with DLM in mind.

\ctrule {FUN-CALL-LIB}
        {\ienvL, \ienvP \vdash f \tk{<{}<{}<} p_1 \tk , p_2 \tk , \dots \tk , p_k \tk{>{}>{}>} \tk ( e_1 \tk , e_2 \tk , \dots \tk , e_n \tk ) \rightarrow_e (\icstr', L')}
        {\ienvL, \ienvP \vdash e_1 \rightarrow_e (\icstr_1, L_{e_1}) \quad e_2 \rightarrow_e (\icstr_2, L_{e_2})
          \dots e_n \rightarrow_e (\icstr_n, L_{e_n})}
        {if $\ienvL \, f = \emptyset$ and \\
          where $L' = \mathlarger\sqcupl_{i=1}^n L_{e_i}$ \\
          and $\icstr' = \bigcup\limits_{i=1}^n \icstr_i$}

The rule below describes how constant labels in the return label of a function are replaced by the label of the expression arguments.
This is done using the $\ienvL'$ environment.
For each parameter that is not defined as a constant (those that have an explicit label) a constraint is added, ensuring that the argument is no more restrictive than the parameter.

\ctrule {FUN-CALL}
        {\ienvL, \ienvP \vdash f \tk{<{}<{}<} p_1 \tk , p_2 \tk , \dots \tk , p_k \tk{>{}>{}>} \tk ( e_1 \tk , e_2 \tk , \dots \tk , e_n \tk ) \rightarrow_e (\icstr_a \cup \icstr_p, L_f')}
        {\begin{gathered}
          \ienvL, \ienvP \vdash e_1 \rightarrow_e (\icstr_1, L_{e_1}) \quad e_2 \rightarrow_e (\icstr_2, L_{e_2}) \quad \dots \quad e_n \rightarrow_e (\icstr_n, L_{e_n}) \\
          \ienvL' \vdash L_f \rightarrow_F L_f'
        \end{gathered}}
        {
          if $\ienvL \, f \neq \emptyset$ and \\
          where $L_f = \ienvL \, f$ \\
          and $((L_1, x_1), (L_2, x_2), \dots, (L_n, x_n)) = \ienvP \, f$ \\
          and $\ienvL' = \ienvL[x_1 \mapsto L_{e_1}, x_2 \mapsto L_{e_2}, \dots, x_n \mapsto L_{e_n}]$ \\
          and $\icstr_a = \bigcup\limits_{i=1}^n \icstr_i$ \\
          and $\icstr_p = \bigcup\limits_{i=1}^n
          \begin{cases}
            \emptyset &\text{if } L_i \text{ is a constant label}\\
            \{ \literal{L_{e_i} \sqsubseteq L_i} \} & \text{otherwise}
          \end{cases}$
          }
