% !TEX root = ../../master.tex

\section{Aiding development of secure software}
When developing any kind of software, developers must take many things into account.
Even when developing software using high abstraction programming languages and/or frameworks, a lot can go wrong when having to take into account the multitudes of aspects related to a problem domain.
If we also add security, specifically the modeling of security policies, even more can go wrong if not properly handled.

Applying more tools and techniques to ensure security can be cumbersome and time-consuming, and could ultimately end up in complete omission if it requires too much effort.
This is why we would like to create a tool which could assist developers in gaining more insights to the security aspects of the software to be written.

\mikkel{More focus on our tool being non-invasive, which is why it is not as time-consuming. Maybe.}
\mikkel{``We imagine a tool that ..... Because it's awesome....''}

We will develop a tool, bottom-up -- as to focus on the practical applications, which will give the user the ability to actively create and modify security policies while implementing software, and receive feedback about breaches of these, in similar style of a modern IDE displaying compiler errors.
\mikkel{Bottom-up in separate sentence, explaining that we take this approach to keep focus on the programmer/us}
This tool will be based on the security policies of the Decentralized Label Model (DLM), with a focus on inferring labels.
We will apply these security policies to the C language.
Additionally, we will introduce our simple time policies, as an extension to the DLM policies, which will be similarly usable and provide similar feedback.

\mikkelin{Why develop for C? Low level language for low level hardware}
