% !TEX root = ../../master.tex

The Internet-of-Things (IoT) is an ever-growing phenomena, becoming apparent in more and more aspects of our everyday lives.\footnote{https://www.theguardian.com/technology/2015/may/06/what-is-the-internet-of-things-google}
However, security in IoT is not widely standardized, and therefore the quality of any such device will depend on the individual manufacturer (as well as any hobbyist making his own devices).
There is still much that can be done in order to expand this area.

In this chapter we will present the needed background information which serves as the base for our project.
We will, based on our pre-specialization \cite{prespecialization}, introduce some of the security problems found in the IoT.
Specifically, we will look into privacy issues associated with IoT.
We will also discuss the possibility of handling some of these privacy issues with time-based constraints.
Additionally, we will discuss some problems related to the development of secure software, which will be related to our solution.
Following the description of our solution, we will present some related work.
Finally, we will present two code examples, which will be used throughout the report, to serve as IoT use-cases.

The remainder of the report is structured as follows.
In \cref{dlm} is provided the necessary background information for understanding the first part of our solution.
Then we will start presenting our solution in \cref{ctif}, by providing an informal description, as well as a formal definition, of how labels are extracted and checked.
Then in \cref{time} this solution is extended, by introducing time policies, which we will compare to existing theory.
Finally, the project is concluded in \cref{conclusion} along with a description of possible future works.
