% !TEX root = ../../master.tex

\section{Security problems}
In \cite{prespecialization} we investigated problems involving the imminent implementation of smart meters throughout Europe.\mikael{Fodnote kilde fra forspeciale.}
Here we discovered a wide array of problems, ranging from ethical to practical.
One of the more interesting discoveries that we made was the newly acquired problems associated with increasing the availability of a previously physically isolated device.
This is exactly what is going to happen when the current electrical meters are going to be substituted with remotely accessed and controlled smart meters.
It will open up for problems already found in similar devices, which in general is any IoT device.

One of the focuses we had in \cite{prespecialization} was on the privacy issues related to such an exposed device.
We would like to explore this problem further, while coming up with a solution that will aid the programmer in the securing IoT devices.
Many of the security problems associated with privacy could possibly also be minimized by adding the concept of time constraints, which we will also discuss here.

\subsection{Privacy}
With the emergence of IoT, more and more of our devices are being globally exposed, for both good and for worse.
There will be endless possibilities in easing our everyday lives, both at a personal level and at a communal level.
At the personal level, we will be able to access devices that monitor and control our homes.
On the communal level, many of the tasks that previously had to be done by people can now be pushed to remote devices, such as monitoring the health of patients or changing traffic signs on a freeway.\footnote{\url{http://techcrunch.com/2015/10/24/why-iot-security-is-so-critical/}}

All these things are associated with major potential risks, should they be compromised.
At the personal level, we do not wish to give up any information about our personal lives, and we don't want anyone else controlling our devices.
At the community level the risks can be even larger, especially those devices linked to health care, such as a ``smart'' pacemaker.

Some of these risks are associated with faulty protocols or erroneous implementations of protocols.
However, privacy can still be an issue even with a ``good'' protocol.
The problem here is that it is not easy ensuring that the information managed by an IoT device is distributed as intended.
This is especially true if different policies apply to different users, as is more than likely with IoT devices.
This could for instance be power consumption data which have different policies depending on who is accessing the data, e.g.: the consumer itself, the electrical company, or the hardware manufacturer.
It is also possible to have \emph{passive cheaters}, users of the system who follow the rules, but try to acquire more information than is intended possible.

\subsection{Time constraints}
Properly handling timing within a system is an already-existing and quite common problem.\footnote{\url{http://www.nist.gov/pml/div688/timing-031915.cfm}}
It is something that is hard to both model and ensure that any such modeling holds.
This is especially true for IoT devices which operate in the real world where any potential mistakes can result in great consequence.

To further the concept of protecting privacy, it would make sense to limit functionality of a device based on time constraints.
By doing this, it could be further ensured that any leak of information from a system is minimal.

\mikkelin{Include \emph{generic} description of how this applies to a password checker and consumption data.
Motivation for the code examples.}